---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# EU Taxonomy Indicator

<!-- badges: start -->
[![R-CMD-check](https://github.com/2DegreesInvesting/tilt.eu.taxonomy/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/2DegreesInvesting/tilt.eu.taxonomy/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of tilt.eu.taxonomy is to assess whether each product of the SME is 
eligible for the EU taxonomy regulation.

## Installation

You can install the development version of tilt.eu.taxonomy from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("2DegreesInvesting/tilt.eu.taxonomy")
```

Or

``` r
# install.packages("pak")
pak::pak("2DegreesInvesting/tilt.eu.taxonomy")
```

## Testing EU Taxonomy Eligibility

The products of each Europages company are matched with the Ecoinvent activities. To test whether Europages products are EU Taxonomy eligible, we will match the Ecoinvent activities of these products with the EU Taxonomy activities. The matching process determines whether the product is taxonomy-eligible or not. If the Ecoinvent activity matches a taxonomy activity, then it is classified as eligible. If the product does not match any taxonomy activity, then it is classified as not eligible.

## Setup

```{r}
library(tilt.eu.taxonomy)
library(dplyr)
options(readr.show_col_types = FALSE)
options(width = 300)
```

## Load Data

### Load Europages products matched with Ecoinvent activities

`ep_companies_ei_activities` is extracted from the dataframe `profile_companies` created by the Python class `PCompanies` in the `tiltIndicatorBefore` repository. It contains Europages companies and products matched with Ecoinvent activities (along with `isic_4digit`). We will select the product-level columns from the whole dataset.

```{r}
ep_products_ei_activities_product <- ep_companies_ei_activities_at_product_level(ep_companies_ei_activities)
ep_products_ei_activities_product |>
  slice(1:5)
```

### EU Taxonomy Activties of EU Taxonomy regulation objectives

EU Taxonomy Regulation sets out six climate and environmental objectives. The six objectives are Climate change mitigation, Climate change adaptation, Protection of water, Transition to a circular economy, Pollution Prevention, and Protection of biodiversity. Each objective outlines those activities which adhere to environmentally sustainable standards. They also outline the information regarding their contribution type, SCC alignment, and DNSH alignment. A subset of activities which follow each objective:

```{r}
# Climate change mitigation
eu_tax_climate_mitigation |>
  slice(11:12)
```

```{r}
# Climate change adaptation
eu_tax_climate_adaptation |>
  slice(1:2)
```

```{r}
# Protection of water
eu_tax_water_protection |>
  slice(2:3)
```

```{r}
# Transition to a Circular Economy
eu_tax_circular_economy_transition |>
  slice(1:2)
```

```{r}
# Pollution Prevention
eu_tax_pollution_prevention |>
  slice(1:2)
```

```{r}
# Protection and restoration of Biodiversity
eu_tax_biodiversity_protection |>
  slice(1:5)
```

### EU Taxonomy Activities' keywords

This prototype focuses on a small subset of 7 taxonomy activities which are shown below. Each taxonomy activity is assigned domain-specific keywords which are chosen manually and with the help of OpenAI's ChatGPT model. They are used to improve the quality of the match between Ecoinevnt activities and EU Taxonomy activities by searching each keyword in the matched ecoinvent activity. Only those EU Taxonomy activities are selected as good matches which have at least one keyword present in the matched ecoinvent activity.

```{r}
eu_tax_activities_keywords
```

### Load NACE-ISIC sector mapper

NACE-ISIC sector mapper serves as the activity matching bridge between the Ecoinvent activities and EU Taxonomy activities.

```{r}
isic_nace_mapper |>
  slice(1:5)
```

## All EU Taxonomy activities with keywords

This step collects all EU Taxonomy activities into a single table from six objective tables. Then, each activity will be assigned domain-specific keywords using `eu_tax_activities_keywords`. We have assigned keywords to only 7 taxonomy activities, hence we will filter out other taxonomy activities in this example.

```{r}
eu_tax_activities_all_with_keywords <- eu_tax_activities_all(
  eu_tax_climate_mitigation, eu_tax_climate_adaptation,
  eu_tax_water_protection, eu_tax_circular_economy_transition,
  eu_tax_pollution_prevention, eu_tax_biodiversity_protection
) |>
  eu_tax_activities_all_with_keywords(eu_tax_activities_keywords) |>
  filter(!is.na(tax_activity_keywords))

eu_tax_activities_all_with_keywords |>
  slice(1:10)
```

## Matching NACE codes to Ecoinvent activities

NACE sector codes are used to match Ecoinvent activities with EU Taxonomy activities. Ecoinvent activities table `ep_products_ei_activities` don't have NACE codes. As a first step, to match NACE codes to `ep_products_ei_activities`, we will use the ISIC sector codes as the mapper present in the `nace_isic_mapper` file.

```{r}
ecoinvent_activities_with_nace <- ecoinvent_activities_with_nace(ep_products_ei_activities_product, isic_nace_mapper)
ecoinvent_activities_with_nace |>
  slice(1:5)
```

## Matching Ecoinvent activities (with NACE) to EU Taxonomy activities

NACE codes are used to match Ecoinvent activities and EU Taxonomy activities. Both the tables `eu_tax_activities_all_with_keywords` and `ecoinvent_activities_with_nace` have the NACE codes. After matching Ecoinvent activities and EU Taxonomy activities, there is a one-to-many relationship between taxonomy activities and Ecoinvent activities respectively. This means that a single taxonomy activity is related to multiple ecoinvent activities.

```{r}
ecoinvent_activities_eu_tax_activities <- ecoinvent_activities_eu_tax_activities(ecoinvent_activities_with_nace, eu_tax_activities_all_with_keywords)
ecoinvent_activities_eu_tax_activities
```

## Keywords search to validate match between Ecoinvent activities and EU Taxonomy activities

As there is a one-to-many relationship between taxonomy activities and Ecoinvent activities, there is a high chance that we might get bad matches between activities. Hence, `taxonomy_keywords` will be used to check whether ecoinvent activities are a closer match to the matched taxonomy activities or not. Each keyword is separated by a semicolon (;), and any value between consecutive semicolons will be matched with ecoinvent activity. If a single taxonomy keyword is present in the `ecoinvent_activity_name` column, the presence is stated with "TRUE" in the `keyword_is_present` column. And "FALSE" is stated if the keyword is not present.

```{r}
eco_activities_after_keywords_search <- eco_activities_after_keywords_search(ecoinvent_activities_eu_tax_activities)
eco_activities_after_keywords_search |>
  slice(1:10)
```

## ChatGPT match validation between Ecoinvent activities and EU Taxonomy activities' description

The EU taxonomy provides a description column for each taxonomy activity which contains more information about them. Hence, we propose to use this additional information to validate whether the ecoinvent activities will be covered by the taxonomy activity description or not. ChatGPT is used to validate the same using `Yes` (covered well) and `No` (not covered well) for each matched `ecoinvent_activity_name` and `tax_activity_description`. We used OpenAI's `gpt-4-turbo` model for validation with the following prompt:

```
"We have a dataframe with columns 'ecoinvent_activity_name' and
'taxonomy_activity_description'. Check row wise whether the ecoinvent activity
(column 'ecoinvent_activity_name') covered by the EU Taxonomy activity description
(column 'taxonomy_activity_description'). Please provide only one word answer
with `Yes` or `No` without any explaination, where `Yes` means covered well and
`No` means not covered well.
"
```
```{r}
eco_activities_after_gpt_validation <- eco_activities_gpt_validation(eco_activities_after_keywords_search)
eco_activities_after_gpt_validation |>
  slice(1:10)
```

## EU Taxonomy-eligible Europages products

The Europages products which have "TRUE" in `keyword_is_present` column and "Yes" in `gpt_validation` column are considered EU Taxonomy-eligible because they are matched with EU Taxonomy activities (via ecoinvent activities) after ISIC-NACE mapping, keywords search, and ChatGPT validation. Here is a list of a small subset of EU Taxonomy-eligible Europages products:

```{r}
tax_eligible_ep_products <- taxonomy_eligible(eco_activities_after_gpt_validation)
tax_eligible_ep_products |>
  slice(30:34)
```

## EU Taxonomy Indicator at company level

After identifying Europages products which are EU Taxonomy-eligible, we want to find out which Europages companies have taxonomy-eligible Europages products. This section loads company level columns and then joins them with matched ecoinvent activities after gpt validation. Lastly, final step filters the Europages companies which have taxonomy-eligible Europages products.

### Load company-level columns from `ep_companies_ei_activities` dataset 

```{r}
ep_products_ei_activities_company <- ep_companies_ei_activities_at_company_level(ep_companies_ei_activities)
ep_products_ei_activities_company |>
  slice(1:10)
```

### Join matched ecoinvent activties after gpt validation to Europages companies 

```{r}
companies_matched_eco_activities <- ep_companies_join_matched_eco_activities(ep_products_ei_activities_company, eco_activities_after_gpt_validation)
companies_matched_eco_activities |>
  slice(1:10)
```

### EU Taxonomy-eligible Europages companies

```{r}
tax_eligible_ep_companies <- taxonomy_eligible(companies_matched_eco_activities)
tax_eligible_ep_companies |>
  slice(1:10)
```
